

# {{< fa user-clock size=1x >}} &nbsp;&nbsp;Exercise (40 min)

## Exercise - Accessing data



**Part 1**: 

* Download the herbarium data "UC1-3c-open-refine.csv", which compiles field observations of plants from Guatemala, available [here](https://docs.gbif.org/course-data-mobilization/exercise-data/). 

  {{< fa lightbulb >}} &nbsp; Use the function [`download.file()`](https://rdrr.io/r/utils/download.file.html) to download the dataset

<br />


## Exercise - Accessing data

**Part 1**: 

* Download the herbarium data "UC1-3c-open-refine.csv", which compiles field observations of plants from Guatemala, available [here](https://docs.gbif.org/course-data-mobilization/exercise-data/). 

  {{< fa lightbulb >}} &nbsp; Use the function [`download.file()`](https://rdrr.io/r/utils/download.file.html) to download the dataset

<br />

* Import the dataset and extract the plant names. How many are there ?

<br />

> We'll focus on 2 species: _Catopsis sessiliflora_ and _Catopsis nutans_. 


## Exercise - Accessing data

**Part 2**: 
<br />

We want to complete the herbarium occurrences of _C. sessiflora_ and _C. nutans_ with extra observations on [GBIF](https://www.gbif.org/). 
<br />

* Download them through a query in R. 
<br />

  {{< fa lightbulb >}} &nbsp; Use the function [`rgbif::occ_search()`](https://docs.ropensci.org/rgbif/reference/occ_search.html) to download species occurrences<br/>



## Exercise - Accessing data

**Part 2**: 
<br />

We want to complete the herbarium occurrences of _C. sessiflora_ and _C. nutans_ with extra observations on [GBIF](https://www.gbif.org/). 
<br />

* Download them through a query in R. 
<br/>

  {{< fa lightbulb >}} &nbsp; Use the function [`rgbif::occ_search()`](https://docs.ropensci.org/rgbif/reference/occ_search.html) to download species occurrences<br/>
<br />

* Inspect the GBIF dataset. What can you see ?
<br />

  {{< fa triangle-exclamation >}} &nbsp; Do not forget to **export the data**

#

## Correction

:::: {.columns}

::: {.column width=45%}

**Part 1**

```{r}
#| echo: true
#| eval: false

## Download herbarium data ----
file_url  <- "https://docs.gbif.org/course-data-mobilization/exercise-data/"
filename <- "UC1-3c-open-refine.csv"

download.file(
  url = paste0(file_url, filename), 
  destfile = "herbarium.csv",
  mode = "wb"
)
```

:::

::: {.column width=55%}
:::

::::


## Correction

:::: {.columns}

::: {.column width=45%}

**Part 1**

```{r}
#| echo: true
#| eval: false

## Download herbarium data ----
file_url  <- "https://docs.gbif.org/course-data-mobilization/exercise-data/"
filename <- "UC1-3c-open-refine.csv"

download.file(
  url = paste0(file_url, filename), 
  destfile = "herbarium.csv",
  mode = "wb"
)

```

<br/>

```{r}
#| echo: true
#| eval: false

## Import dataset ----
df <- read.csv2("herbarium.csv")

## Have a look ----
df |> head()

```

:::

::: {.column width=55%}
:::

::::



## Correction


:::: {.columns}

::: {.column width=45%}

**Part 1**

```{r}
#| echo: true
#| eval: false

## Download herbarium data ----
file_url  <- "https://docs.gbif.org/course-data-mobilization/exercise-data/"
filename <- "UC1-3c-open-refine.csv"

download.file(
  url = paste0(file_url, filename), 
  destfile = "herbarium.csv",
  mode = "wb"
)

```

<br/>

```{r}
#| echo: true
#| eval: false

## Import dataset ----
df <- read.csv2("herbarium.csv")

## Have a look ----
df |> head()

```
:::

::: {.column width=5%}
:::


::: {.column width=45%}

<br/><br/>

```{r}
#| echo: true
#| eval: false

## Extract plant names ----
df$Full.name |> unique()
```

<br/>

We'll focus on 2 species: _Catopsis sessiliflora_ and _Catopsis nutans_. 

<br/>

```{r}
#| echo: true
#| eval: false

## Extract plant names ----
plant_species <- c("Catopsis sessiliflora", "Catopsis nutans")

```


:::


::::

## Correction


:::: {.columns}

::: {.column width=45%}

**Part 2**

<br/>

```{r}
#| echo: true
#| eval: false

## Download GBIF occurrences ----
occ <- rgbif::occ_search(
  scientificName = plant_species, 
  fields = c('name', 'scientificName' , 'decimalLatitude', 'decimalLongitude', 'country'), 
  country = "GT",
  hasCoordinate = TRUE,
  hasGeospatialIssue = FALSE
)

## Append occurrences -----
occ <- rbind(
  occ$`Catopsis sessiliflora`$data,
  occ$`Catopsis nutans`$data
)

## Export ----
write.csv2(
  x = occ,
  file = "gbif_occurrences.csv", 
  row.names = FALSE
)

```
:::

::: {.column width=5%}
:::



::: {.column width=45%}


<br/><br/><br/>

```{r}
#| echo: true
#| eval: false

# Import GBIF occurrences ----
occ <- read.csv2("gbif_occurrences.csv")

## Inspect ----
head(occ)

```  
:::

We will need to homogenize the herbarium data and GBIF data before we can pool the data together ! 

::::



