# {{< fa code size=1x >}} &nbsp;&nbsp;Scripting (10 min)

<!-- Side 1 -->
## Queries from R

R lets you download and import the data directly if you have the url, by using the `download.file()` function:

```{r}
#| echo: true
#| eval: false

download.file("https://geodata.ucdavis.edu/gadm/gadm4.1/gpkg/gadm41_FRA.gpkg", destfile = here::here("data", "fr_2.gpkg"))
```



<!-- Side 1 -->
## Queries from R

R lets you download and import the data directly if you have the url, by using the `download.file()` function:

```{r}
#| echo: true
#| eval: false

download.file("https://geodata.ucdavis.edu/gadm/gadm4.1/gpkg/gadm41_FRA.gpkg", destfile = here::here("data", "fr_2.gpkg"))
```

> Sometimes it sends strange errors on windows. Simply add  `mode = "wb"`

```{r}
#| echo: true
#| eval: false

download.file("https://geodata.ucdavis.edu/gadm/gadm4.1/gpkg/gadm41_FRA.gpkg", destfile = here::here("data", "fr_2.gpkg"), mode = "wb")
```



<!-- Side 2 -->

## What if the file is compressed ?

You can use the helper functions `unzip()`, `gzfile()` etc but it's unnecessary for most R reading functions:

```{r}
#| echo: true
#| eval: false

download.file("http://www.sociopatterns.org/wp-content/uploads/2015/07/Friendship-network_data_2013.csv.gz", destfile = here::here("data", "friends.gz"))

# all of these work fine: 
friends <- read.table(gzfile( here::here("data", "friends.gz")))
friends <- read.csv2(gzfile( here::here("data", "friends.gz")), sep = " ")
friends <- read.table(here::here("data", "friends.gz"))
```


<!-- Side 2 -->

## What if the file is compressed ?

You can use the helper functions `unzip()`, `gzfile()` etc but it's unnecessary for most R reading functions:

```{r}
#| echo: true
#| eval: false

download.file("http://www.sociopatterns.org/wp-content/uploads/2015/07/Friendship-network_data_2013.csv.gz", destfile = here::here("data", "friends.gz"))

# all of these work fine: 
friends <- read.table(gzfile( here::here("data", "friends.gz")))
friends <- read.csv2(gzfile( here::here("data", "friends.gz")), sep = " ")
friends <- read.table(here::here("data", "friends.gz"))
```

Other useful packages: `data.table`, `vroom`, ...



